Quality of Service with Software Defined Network.

### 2017-03-05 report/paper.tex
corrected some parts written by matteo and added some other parts

\section{Needs and Gools}\label{sec:obb} %obbiettivi ed esigenze controllare
The main gool of this experience is to develop a small SDN network in a lab with real hardware
in orther to understood come a livello pratico si possa implementare  a QoS service in a SDN scenarois.
Our network have to manage traffic with a global scope, in order to evolve in base on different parameters.
Inoltre the SDN network need to interact with world wide web efficently.
%correggere
  \subsection{Adaptability of the network}
  Made Network Adaptable means poter dirottare il traffico in real time to ensure a lot of demands. For example we have to avoid link congestion,
  but we can prefer reroute some hosting traffic to peggiorare it's performance because of it's evil traffic.
  To reach this gools we need to have a global knowledge on the network and we have to modify the routing table on every routerboard with a
  global scope.
  The performance of a single connection and entire network can ensure various type of metrics.
  In order to assicurare queste ci serve una struttura dati che si aggiorna costantemente nella quale immagazzinare le statistiche dei link
  in modo all'occorrenza di trovare il percordo migliore per ogni singolo flusso dati.
  \#TODO MATTEO
  \subsection{Exterior comunication}
  \#TODO pitt

@@

  \subsection{Traffics analisis}
  For every networks performance we have tryed to develop a specific solution. The OpenFlow main goal is to centralize controll and software
  complexity so the routerboard can made only some easy function in orther to contain the hardware complexity and mantainly a good speed.
  Il che ha portato a dover utilizzare stratagemmi.
  \#TODO %riformulare
    \subsubsection{Delay}
    Per molte applicazioni odierne il delay e' un parametro importantissimo.
    La rete attuale puo' conoscere il delay di un link semplicemente inviando un pachetto ICMP di tipo TimeStamp ( verificare )
    supportato dal protocollo openflow, questo permette di ottenere una precisione nell'ordine dei microsecondi ( verificare ).

    \#TODO MATTEO
		\subsubsection{Link Capacity}
		\#TODO andrea
		The first data that was collected was the actual link capacity, which depends both on the cable used to connect
		different devices and the effective capacity for each ethernet port linked by the cable itself. Everytime a device supporting
		openflow connects to the controller, it gives to the latter information about its status and its capabilities. One of them is
		the link capacity in three ways: the first is the \textit{supported} capacity, which is the phisical maximum capacity, the second is the
		\textit{advised} capacity, which name describes itself and the latter is the \textit{current} one. We choose to save the last one,
		as it gives the most useful about the link state. This information, moreover, is the one that non-SDN switches use to choose the \textit{metric}
		parameter. The simulation with Mininet showed us that this type of information can be used, but the Routeboards didn't support it.
		As a matter of fact, we weren't able to exploit this capacity in our network realization.
    \subsubsection{Throughput}
    \#TODO andrea
		Using the flow statistics, we analyzed the traffic going through a link and, with a simple division, we got the normalized throughput.
		We exploited the information given by the switches with the match structure: this means that we used the throughput for the TCP traffic,
		the one for the UDP and for each of them we could know toward which IP the traffic is directed or is from. This was exploited with mininet
		simulations, but not in the set up in the laboratory. With the latter, we used the statistics as an indicator of the bitrate of a link
		and divided the host to categories based on it.

		\subsubsection{Packet error rate}
		The port statistics given by the switches contains different informations about the state of each port in a switch. To classify a link for its
		reliability, we divided the number of packets with erroneous symbols that couldn't be corrected with a FEC by the total number of packets that were sent,
		getting the packet error rate. The choice we made when exploiting this statistic was privileging the gaming and normal traffic, wherease the
		heavy traffic was directed to links with highest packet error rate.
		\#TODO andrea

@@

  \subsection{Route changing} %cambio rotte
  \#TODO andrea
	For each statistic we got and each parameter we obtained, we changed the weight of the graphs. Every graph we used had a specific weight, so the
	Dijkstra algorithm for a weighted graph gave different minimum path between two hosts. From the parameter we saved, we choose to change the
	route to reach the destination and viceversa. The traffic follow the same path, in both the directions and depends only on the type of traffic:
	we choose to set the minimum unweighted path if no statistics were available, the minimum throughput path for the standard traffic, the minimum
	delay for the gaming one and the maximum packet error rate for the heavy traffic generated by a host.
	The changing of the path were made only if the last change happened before one minute because the torrent traffic has, in its protocol, a discovery
	algortithm. This means that if a lot of changes were made, most of the traffic generated by the peer to peer, is the control one and the result is that the
	load in the network will be heavy and last longer.

  \subsection{ARP solution} %TODO questione dell'arp inutile
  During our analysis and in our tests, we prove that in a SDN mesh network, the ARP packets are useless.
  This type of packets is useful for a multipoint to multipoint connection where the connection is set between a
  HUB and a set of computers or between a bridge and a computer. The first case is no more in use in common networks as
  hiher the collision range. The bridges limits their collision range to the port. With openflow each switching unit can
  act as a bridge or as a router. In both cases analyses the header of the L2 packet L3 packet and some of the L4 header.
  If it has the rule to forward the packet, it will act as declared in the most specific matching rule. Otherwise it will
  send to the controller a PacketIn and the controller will act as programmed. If an host want to communicate to another host whithin
  the network, it sends the packet to the line and then the switch will check in its tables and execute the actions of the match found.
  For an IP network, it will check the path to reach a destination and then send to the defined output port. The mac address in this case
  became useless, as the connection is point to point. So, knowing the destination layer2 address is useless as the switch can act
  regardless of it. The only thing the switch need to know is the layer2, layer 3 addresses and, at most, the layer 4 address.
  This can be seen by a reader as a violation of the layers in the ISO/OSI stack, but to exploit the potentials of the SDN approach,
  this is necessary. TODO explain why TODO!!!!**
  \#TODO andrea

  \subsection{Default Gw problem}
  The computers don't need anymore to set a gateway to reach another network as the controller and the switches will redirect the traffic
  properly, according to the rules set by the first. The controller must only know which switch ports are connected to the
  internet and then add them to the path. in this way, the controller can set a path to reach it and provide connectivity with
  a traffic shaping policy and/or differentiate the traffic from and to different networks.
  \#TODO andrea

  \subsection{Ip switching in pox}
  \#TODO andrea
	One of the problem encountered was that, if an host is already found by pox and there is a PacketIn event, the core of the controller
	switches the source and the destination IP and MAC address with the source one. The problem was partially solved, checking out the
	parameters known by the controller, i.e. if the two host are connected to two different switches and the controller switches the source with the
	destination address, we can switch it back; the result is that the PacketIn event now has correct informations.

>>>>

\section{Needs and Goals}\label{sec:obb} %obbiettivi ed esigenze controllare
The main gool of this experience is to develop a small SDN network in a lab with real hardware
in orther to understand in a practical level how to implement a \textit{Quality of Service, \textbf{QoS}} service in a SDN scenarios.
Our network have to manage traffic with a global scope, in order to evolve in base on different parameters.
Moreover the SDN network need to interact with world wide web efficently.
%correggere
  \subsection{Adaptability of the network}
  Made Network Adaptable means poter dirottare il traffico in real time to ensure a lot of demands. For example we have to avoid link congestion,
  but we can prefer reroute some hosting traffic to worsen it's performance because  it is the traffic we don't want to encourage.
  To reach this gools we need to have a global knowledge on the network and we have to modify the routing table on every routerboard with a
  global scope.
  The performance of a single connection and entire network can ensure various type of metrics.
  In order to ensure the traffic shaping based on network informations, there is a need of a structure, automatically updatable, where the link statistics
	can be saved to allow the best path for a data stream. The best metric will be described below.
	%  assicurare queste ci serve una struttura dati che si aggiorna costantemente nella quale immagazzinare le statistiche dei link
  % in modo all'occorrenza di trovare il percordo migliore per ogni singolo flusso dati.
  \#TODO MATTEO
  \subsection{Exterior comunication}
  \#TODO pitt
	Our network needed to be able to communicate with the external internet to be useful for a demo which demonstrate the capabilities of a SDN network.
	The first thing we needed to accomplish was to make the controller aware of what switch and which port of the switch is connected to the external internet:
	to do so, we allowed our controller to be able to receive in-line parameters. The two parameters we needed the controller to be aware were the mac address of the
	port of a switch that is connected to the link and the network address of our network, to let the controller know if a request from a host was meant to another host
	within the network or the data stream was intended to the internet.
	In this way the host connected to the inner network can surf the net regardless of which type of devices are switching and routing the traffic they are generating.

@@

  \subsection{Traffics analisis}\label{subsec:traf}
  For every networks performance we have tryed to develop a specific solution. The OpenFlow main goal is to centralize controll and software
  complexity so the routerboard can made only some easy function in orther to contain the hardware complexity and mantainly a good speed.
  Il che ha portato a dover utilizzare stratagemmi.
  \#TODO %riformulare
    \subsubsection{Delay}

	  Per molte applicazioni odierne il delay e' un parametro importantissimo.
    La rete attuale puo' conoscere il delay di un link semplicemente inviando un pachetto ICMP di tipo TimeStamp ( verificare )
    supportato dal protocollo openflow, questo permette di ottenere una precisione nell'ordine dei microsecondi ( verificare ).

    \#TODO MATTEO
		\subsubsection{Link Capacity}

		\#TODO andrea
		The first data that was collected was the actual link capacity, which depends both on the cable used to connect
		different devices and the effective capacity for each ethernet port linked by the cable itself. Everytime a device supporting
		openflow connects to the controller, it gives to the latter information about its status and its capabilities. One of them is
		the link capacity in three ways: the first is the \textit{supported} capacity, which is the phisical maximum capacity, the second is the
		\textit{advised} capacity, which name describes itself and the latter is the \textit{current} one. We choose to save the last one,
		as it gives the most useful about the link state. This information, moreover, is the one that non-SDN switches use to choose the \textit{metric}
		parameter. The simulation with Mininet showed us that this type of information can be used, but the Routeboards didn't support it.
		As a matter of fact, we weren't able to exploit this capacity in our network realization.
    \subsubsection{Throughput}

	  \#TODO andrea
		Using the flow statistics, we analyzed the traffic going through a link and, with a simple division, we got the normalized throughput.
		We exploited the information given by the switches with the match structure: this means that we used the throughput for the TCP traffic,
		the one for the UDP and for each of them we could know toward which IP the traffic is directed or is from. This was exploited with mininet
		simulations, but not in the set up in the laboratory. With the latter, we used the statistics as an indicator of the bitrate of a link
		and divided the host to categories based on it.

		\subsubsection{Packet error rate}

		The port statistics given by the switches contains different informations about the state of each port in a switch. To classify a link for its
		reliability, we divided the number of packets with erroneous symbols that couldn't be corrected with a FEC by the total number of packets that were sent,
		getting the packet error rate. The choice we made when exploiting this statistic was privileging the gaming and normal traffic, wherease the
		heavy traffic was directed to links with highest packet error rate.  As the switches exchanges a packet before recognizing most of the error, both because the
		phisical error are corrected using blocks coding and other error are found using a \textit{Frame Check Sequence} like CRC, the byte error were considered worse as
		those are indirect measures themself and the error derived from the calculations could grow up.
		\#TODO andrea

@@

  \subsection{Route changing} %cambio rotte

  \#TODO andrea
	For each statistic we got and each parameter we obtained, we changed the weight of the graphs. Every graph we used had a specific weight, so the
	Dijkstra algorithm for a weighted graph gave different minimum path between two hosts. From the parameter we saved, we choose to change the
	route to reach the destination and viceversa. The traffic follow the same path, in both the directions and depends only on the type of traffic:
	we choose to set the minimum unweighted path if no statistics were available, the minimum throughput path for the standard traffic, the minimum
	delay for the gaming one and the maximum packet error rate for the heavy traffic generated by a host.
	Two threshold were chosen based on the traffic analysis described on Sec. \ref{subsec:traf} which are 200 bytes/s for the gaming traffic and one megabyte/s
	when the node is considered in heavy traffic. Those two threshold were made as they are average values and considering that our network is made of few really fast link,
	so on one side the lower limit is sufficient not to consider gaming traffic to be a normal traffic, wherease it's easy to reach the high level of traffic load between two machines.
	The changing of the path were made only if the last change happened before one minute because the torrent traffic has, in its protocol, a discovery
	algortithm. This means that if a lot of changes were made, most of the traffic generated by the peer to peer, is the control one and the result is that the
	load in the network will be heavy and last longer.

  \subsection{ARP solution} %TODO questione dell'arp inutile

  During our analysis and in our tests, we prove that in a SDN mesh network, the ARP packets are useless.
  This type of packets is useful for a multipoint to multipoint connection where the connection is set between a
  HUB and a set of computers or between a bridge and a computer. The first case is no more in use in common networks as
  hiher the collision range. The bridges limits their collision range to the port. With openflow each switching unit can
  act as a bridge or as a router. In both cases analyses the header of the L2 packet L3 packet and some of the L4 header.
  If it has the rule to forward the packet, it will act as declared in the most specific matching rule. Otherwise it will
  send to the controller a PacketIn and the controller will act as programmed. If an host want to communicate to another host whithin
  the network, it sends the packet to the line and then the switch will check in its tables and execute the actions of the match found.
  For an IP network, it will check the path to reach a destination and then send to the defined output port. The mac address in this case
  became useless, as the connection is point to point. So, knowing the destination layer2 address is useless as the switch can act
  regardless of it. The only thing the switch need to know is the layer2, layer 3 addresses and, at most, the layer 4 address.
  This can be seen by a reader as a violation of the layers in the ISO/OSI stack, but to exploit the potentials of the SDN approach,
  this is necessary. TODO explain why TODO!!!!**
  \#TODO andrea

  \subsection{Default Gw problem}

  The computers don't need anymore to set a gateway to reach another network as the controller and the switches will redirect the traffic
  properly, according to the rules set by the first. The controller must only know which switch ports are connected to the
  internet and then add them to the path. in this way, the controller can set a path to reach it and provide connectivity with
  a traffic shaping policy and/or differentiate the traffic from and to different networks.
  \#TODO andrea

  \subsection{Ip switching in pox}

  \#TODO andrea
	One of the problem encountered was that, if an host is already found by pox and there is a PacketIn event, the core of the controller
	switches the source and the destination IP and MAC address with the source one. The problem was partially solved, checking out the
	parameters known by the controller, i.e. if the two host are connected to two different switches and the controller switches the source with the
	destination address, we can switch it back; the result is that the PacketIn event now has correct informations.
